Formalizing Stack Safety as a Security Property

### 2022-12-07 submission5/core.tex
wordsmith; add compatiblity terminology

To characterize this situation in a property, we use the concept of {\it non-interference}
based on variant states. Specifically, we consider two values to be variants, even if their
payload values differ, as long as the remainder of their state respects other system-specific
constraints; in the case of the tagged system, the tag component must be identical.
Then, for a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
they agree on the value of every element not in \(\components\), and the values
in \(\components\) are variants.

We say that the elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if every \(\components\)-variant state produces an equivalent observation trace
(see Section \ref{sec:events}); all other elements are \emph{relevant}.
We characterize \emph{observational integrity} as the property that
every relevant element that is \(\sealed\) under the callee's view is restored
to its original value at the return point.

@@

\definition For a given system, assume an equivalence relation \(\sim\) on values.
Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if, for
all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\),
and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).

>>>>

To characterize this situation in a property, we borrow the idea of \emph{variant} states
from the theory of non-interference. 
For a set of elements \(\components\),
a pair of states \(\mach\) and \(\nach\) are {\em \(\components\)-variants} if
their values an only disagree on elements in \(\components\). We further require that
the respective values in \(\components\) be \emph{compatible}, meaning that even though
their payloads may differ, their tags (or whatever other data is important to
the machine's security enforcement mechanism) must be identical.

We say that the elements of \(\components\) are \emph{irrelevant}
in \(\mach\) if they can be replaced by compatible values without changing the
observable behavior of the machine. (We will formalize this in Section~\ref{sec:events}
by saying that every \(\components\)-variant state produces
an equivalent observation trace.) All other elements are \emph{relevant}.
We characterize \emph{observational integrity} as the property that
every relevant element that is \(\sealed\) under the callee's view is restored
to its original value at the return point.

@@

For a given system, assume an compatibility equivalence relation \(\sim\) on values.
Two states are variants with respect to a set of elements \(\components\)
if they agree on the value of every element not in \(\components\) and
and have compatible values for every element in \(\components\).
Our notion of non-interference involves comparing the traces of such
\(\components\)-variants. We use this to define sets of irrelevant elements,
and secrets in general.

\definition Machine states \(\mach\) and \(\nach\) are {\em \(\components\)-variants},
written \(\mach \approx_\components \nach\), if, for
all \(\component \not \in \components\), \(\mach[\component] = \nach[\component]\)
and for all \(\component \in \components\), \(\mach[\component] \sim \nach[\component]\).
